\lett{T}{he work described} in this dissertation 
could not have been completed 
without the assistance and support from  
various people and institutions over the years, 
and it is a pleasure to take this opportunity 
to pay tribute to those who have aided in its completion.


%Schols, SFI, AMBER, RIA
%%%%%%%%%%%%%
I would first like to thank 
Science Foundation Ireland 
for providing the funding for this research   
through the {\sc AMBER} program, 
as well as the Royal Irish Academy 
for facilitating the project collaboration with 
our colleagues at 
the University of Liverpool.
%
I also wish to thank 
Trinity College for their generous support 
through the provision of 
the Non-Foundation Scholarship.
%which generously provided my 
%accommodation and tuition fees.  
%during the course of my Ph.D..
%
I am grateful to the staff at the School of Physics 
for their assistance over the years, 
and to Trinity Research IT and {\sc ICHEC} 
for providing computing resources, 
as well as the 
{\sc CRANN} and {\sc AMBER} support staff.
%%%%%%%%%%%%%


%Collaborators - Johnny + co, Emma + co, ONETEP
%%%%%%%%%%%%%
Much of this work would not be possible 
were it not for my esteemed colleagues.
%
%Spintronics people
%%%%%%%%%%%%%
I would like to thank the members of the 
Computational Spintronics Group 
for their constructive feedback 
and discussions when presented with my work.
% 
I am deeply grateful in particular to 
my co-supervisor, 
Professor Stefano Sanvito, 
for all of his valuable help 
and guidance in my studies.

The members of the 
Chemical Physics of Low-Dimensional 
Nanostructures Group, 
Professor Johnny Coleman, 
Dr Damien Hanlon, 
Dr Conor Boland, 
and Dr Claudia Backes 
deserve a special thanks 
for expediting my first publication.
%
I am grateful also to 
Emma Norton and Dr Karsten Fleischer 
of the Applied Physics Research Group
for their helpful discussions and 
for contributing their experimental data.
%and Professor Cormac McGuinness, 
%
I also appreciate the helpful discussions with 
Fiona McCarthy, Mark McGrath,  
Thomas Wyse Jackson, 
and Edward Linscott.
%%%%%%%%%%%%%


%ONETEP 
%%%%%%%%%%%%%
The members of the {\sc ONETEP} community 
were very helpful to me during the coding retreats. 
I appreciate their assistance very much.
%%%%%%%%%%%%%

%Gilberto
%%%%%%%%%%%%%
My time in Dublin 
was frequently punctuated by 
visits to the University of Liverpool 
and I was incredibly fortunate to be placed 
in a safe pair of hands upon arrival.
%
I would therefore like to extend 
my warmest appreciation   
to the venerable Dr Gilberto Teobaldi 
for making each of these visits 
an absolute pleasure, 
even on those dark December days.
%
His valuable assistance, 
inexhaustible patience, 
and excellent taste in hotels  
aided my stay immeasurably.
%
Many thanks + cheer up.
%%%%%%%%%%%%%


%OFFICE LADS
%%%%%%%%%%%%%
I would like to give a special 
mention to my office-mates, 
who I have had the 
utmost pleasure of working beside:
Dr James Lawlor, Dr Eric Mehes, 
Dr Pankaj Kumar, Dr Alessandro Lungi, 
Dr Claudia Rocha, Dr Linda Zotti, 
and last, but certainly not least, 
Okan Karaca Orhan.
%
You have each made a valuable 
contribution to my postgraduate experience.
%
I thank you all for the welcome distractions.
%%%%%%%%%%%%%


%Izzy, Naomi, Lauren, Niamh
%%%%%%%%%%%%%
It would be remiss of me to fail to mention 
my distinguished flatmates, 
Izzy Kennedy, Lauren Boland, 
Naomi Beard and Niamh Ennis,
whose company was the 
highlight of my years living on 
Trinity College campus.
%


%David
%%%%%%%%%%%%%
I would like to extend my heartfelt gratitude 
to my supervisor, Professor David O'Regan, 
to whom I am truly indebted most of all, 
not least for taking a fellow Corkman under his wing.
%
I have been extremely fortunate 
to be under his considerate mentorship 
the last four years.
%
His continuous encouragement, 
attentive guidance, 
and meticulous work ethic 
are an example to aspire to.
%%%%%%%%%%%%%


%Olivia
%%%%%%%%%%%%%
I am especially privileged 
to have had Olivia Kerrigan 
by my side the last few years.
%
I can only hope to repay 
the patience and devotion 
she has afforded me during 
these months of writing.
%
She continues to brighten  
the otherwise anarchic days 
of the Ph.D. experience. 
%
I also humbly thank 
the Kerrigan family for 
graciously hosting me 
for the final push.
%%%%%%%%%%%%%

%Friends + Family
%%%%%%%%%%%%%
Finally, 
I would to thank my friends and family 
who have made my time here in Dublin 
over the past eight years very special.
%
In particular, 
my parents, Karen and Michael, 
grandparents, and siblings 
who have all supported me relentlessly  
in the Big Smoke. 
%
I wouldn't be here without them.
%%%%%%%%%%%%%





